Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I

### 2025-05-20 manuscript/manuscript.Rmd
Fixing subscripts

Other amino acids and nucleobases/nucleosides had trace amounts of labeling after 3 hours and were extensively labeled (>25%) after 24 hours (Figure 1). Cytidine (42.2% labeled) and adenosine (59.9%) were more heavily ^15^N labeled than their nucleobase constituents at T26 (22.9% and 42.1%, respectively), corresponding to the expected synthesis pathway. Guanine, however, had the inverse trend and at 44.9% labeled was 15% more labeled than guanosine (Supplemental figure XX). The deoxyribose forms were in all cases less completely labeled than their ribose counterparts (deoxycytidine = 18.9%, deoxyadenosine = 20.3%, and deoxyguanosine = 17.2%) `r wc("Maybe better as a small table?")` `r wc("Consider trying harder to get thymine/uracil/inosine values?")`. In each case, the fully ^15^N labeled nucleobase was the dominant constituent of the isotope envelope (^15^N_3_ for cytosine and ^15^N_5_ for adenine and guanine). Other compounds, including betaines (here restricted to quaternary amines with a carboxylic acid group) and choline derivatives showed very little labeling at all, with only homarine (15.0% labeled) and O-acetylcarnitine (9.9% labeled) exceeding 5% labeling after 24 hours`r wc("Expand betaine discussion and turn it into a separate paragraph? Not sure what more to say...")`.

Amendments with ^15^NO_3_ showed a significantly different labeling pattern with very few metabolites incorporating the ^15^N label at the surface even after 24 hours. Glutamine and glutamate, which were labeled within 20 minutes from ammonia and are the main way in which inorganic nitrogen becomes organic, here saw only trace amounts of production from the added nitrate even after 24 hours. The most heavily labeled compound was guanine, which saw trace amounts of labeling at the T10 timepoint and reached ~40% labeling at T26. This labeling was not reflected in the other nucleobases. The only group of compounds that did see compelling degrees of label were the mycosporine-like amino acids (MAAs), a group of compounds produced under UV exposure (see below). These compounds, like guanine, were labeled in excess of the bulk glutamate/amino acid pool from which their nitrogen is sourced. This is only possible if the nitrate is assimilated by a subset of the community that is able to reduce it and prevents it from entering the environment. Metabolites specific to those organisms can then be made from the small pool of heavily-labeled organic material without affecting the fraction of glutamate or glutamine labeled in bulk.

@@

The organic nitrogen amendments also showed labeling patterns distinct from the ammonia incubations, though a large fraction of the nitrogen was clearly remineralized into ammonia/glutamate and used to label a similar suite of compounds (Figure 1). Fully C and N labeled (^13^C_6_, ^15^N_4_) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours. Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints showing

>>>>

Other amino acids and nucleobases/nucleosides had trace amounts of labeling after 3 hours and were extensively labeled (>25%) after 24 hours (Figure 1). Cytidine (42.2% labeled) and adenosine (59.9%) were more heavily ^15^N labeled than their nucleobase constituents at T26 (22.9% and 42.1%, respectively), corresponding to the expected synthesis pathway. Guanine, however, had the inverse trend and at 44.9% labeled was 15% more labeled than guanosine (Supplemental figure XX). The deoxyribose forms were in all cases less completely labeled than their ribose counterparts (deoxycytidine = 18.9%, deoxyadenosine = 20.3%, and deoxyguanosine = 17.2%) `r wc("Maybe better as a small table?")` `r wc("Consider trying harder to get thymine/uracil/inosine values?")`. In each case, the fully ^15^N labeled nucleobase was the dominant constituent of the isotope envelope (^15^N~3~ for cytosine and ^15^N~5~ for adenine and guanine). Other compounds, including betaines (here restricted to quaternary amines with a carboxylic acid group) and choline derivatives showed very little labeling at all, with only homarine (15.0% labeled) and O-acetylcarnitine (9.9% labeled) exceeding 5% labeling after 24 hours`r wc("Expand betaine discussion and turn it into a separate paragraph? Not sure what more to say...")`.

Amendments with ^15^NO~3~ showed a significantly different labeling pattern with very few metabolites incorporating the ^15^N label at the surface even after 24 hours. Glutamine and glutamate, which were labeled within 20 minutes from ammonia and are the main way in which inorganic nitrogen becomes organic, here saw only trace amounts of production from the added nitrate even after 24 hours. The most heavily labeled compound was guanine, which saw trace amounts of labeling at the T10 timepoint and reached ~40% labeling at T26. This labeling was not reflected in the other nucleobases. The only group of compounds that did see compelling degrees of label were the mycosporine-like amino acids (MAAs), a group of compounds produced under UV exposure (see below). These compounds, like guanine, were labeled in excess of the bulk glutamate/amino acid pool from which their nitrogen is sourced. This is only possible if the nitrate is assimilated by a subset of the community that is able to reduce it and prevents it from entering the environment. Metabolites specific to those organisms can then be made from the small pool of heavily-labeled organic material without affecting the fraction of glutamate or glutamine labeled in bulk.

@@

The organic nitrogen amendments also showed labeling patterns distinct from the ammonia incubations, though a large fraction of the nitrogen was clearly remineralized into ammonia/glutamate and used to label a similar suite of compounds (Figure 1). Fully C and N labeled (^13^C~6~, ^15^N~4~) arginine was visibly taken up at the T0 timepoint (within 20 minutes) and 80-90% of the pool was fully labeled after 12 hours. Surprisingly, the fraction of arginine labeled decreased from this point, with the T26 and T73 timepoints showing
